### 2019-01-24 Paper/Untitled/Untitled.Rmd

---
title: 'A template for writing manuscripts in Rmarkdown'

author:

- name: Jeffrey W. Hollister
  affilnum: '1,2'

- name: Francisco Rodriguez-Sanchez 
  affilnum: 3
  email: f.rodriguez.sanc@gmail.com

affiliation:

- affilnum: 1
  affil: US Environmental Protection Agency, Office of Research and Development, National Health and Environmental Effects Research Laboratory, Atlantic Ecology Division, 27 Tarzwell Drive  Narragansett, RI, 02882, USA

- affilnum: 2
  affil: Big Name University, Department of R, City, BN, 01020, USA

- affilnum: 3
  affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain

output:

  pdf_document:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
    includes:
      in_header: header.tex

  html_document: null

  word_document: null

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: doublespacing
footerdate: yes
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

---

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")

```

```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

\singlespace

\vspace{2mm}\hrule

Write your abstract here.

\vspace{3mm}\hrule

*Keywords*: rmarkdown, reproducible science

\doublespace

\bleft

# INTRODUCTION

Write your introduction here. You can cite bibliography like this [@Yan2011; @Sutherland2011], if you provide a `BibTeX` file with references. See http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html for more information. Or you could also use [knitcitations](https://cran.r-project.org/web/packages/knitcitations/index.html) or [RefManageR](https://cran.r-project.org/web/packages/RefManageR/index.html) to fetch bibliographic metadata automatically from the web. For example, citing a paper can be as easy as providing its DOI `r citep("10.1016/j.tree.2006.03.016")` or even just a few keywords `r citep("Ricklefs 2008 American Naturalist")`. They will then automagically appear in the list of cited references.

You can even specifiy the desired output format for your bibliography by including a style file for a specific journal (e.g. "ecology.csl"). Many different bibliography styles (CSL files) can be obtained at http://citationstyles.org/ or https://github.com/citation-style-language/styles.

# METHODS

## Study Area

We worked in a **beautiful** place with lots of trees, like *Quercus suber* and *Laurus nobilis*.

## Data collection and analysis

```{r read_data, eval = FALSE, echo = FALSE}
dataset <- read.csv("mydata.csv")
```

We applied a linear model where

$$
y_{i} = \alpha + \beta*x_{i} 
$$

```{r model, echo = FALSE, eval = FALSE}
model <- lm(y ~ x)
```

We used the statistical language `R` `r citep(citation())` for all our analyses. These were implemented in dynamic rmarkdown documents using `knitr` `r citep(citation("knitr"))` and `rmarkdown` `r citep(citation("rmarkdown"))` packages. All the multilevel models were fitted with `lme4` `r citep(citation("lme4"))`.

# RESULTS

Trees in forest A grew taller than those in forest B (mean height: `r mean(25, 31, 28)` versus `r mean(13, 19, 16)` m). And many more cool results that get updated dynamically.

# DISCUSSION

Discuss.

# CONCLUSIONS

# ACKNOWLEDGEMENTS

# REFERENCES

```{r write_citations, cache=FALSE, include=FALSE}
write.bibtex(file="knitcitations.bib")
```

<div id = "refs"></div>

\eleft

\clearpage

\listoftables

\newpage

```{r Table1, results='asis', echo=FALSE, cache=FALSE}
kable(head(iris), caption = "A glimpse of the famous *Iris* dataset.")
```

\newpage

```{r Table2, results='asis', echo=FALSE, cache=FALSE}
kable(mtcars[10:16, ], caption = "Now a subset of mtcars dataset.")
```

\clearpage

\listoffigures

\newpage

```{r Fig1, echo=FALSE, fig.cap="Just my first figure with a very fantastic caption.", cache=FALSE}
x <- rnorm(100)
y <- jitter(x, 1000)
plot(x, y)
```

\newpage

\blandscape

```{r Fig2, echo=FALSE, fig.cap="Second figure in landscape format.", cache=FALSE}
a <- sort(rnorm(100))
b <- c(rep("Group Small", 35), rep("Group Big", 65))
boxplot(a ~ b)
```

\elandscape

\clearpage

```{r sessioninfo, echo = FALSE, eval = FALSE}
# set eval = FALSE if you don't want this info (useful for reproducibility) to appear 
sessionInfo()
```