### 2019-01-24 Paper/AgreementAttraction.Rmd
paperwriting

---
title: 'Agreement Attraction in Turkish: Effects of Nominal and Verbal Plural Morphemes'

author:

- name: Utku Turk
  affilnum: 1
  email: utku.turk@boun.edu.tr

- name: Pavel Logacev 
  affilnum: 1

affiliation:

- affilnum: 1
  affil: Bogazici University, South Campus, John Freely Building, Department of Linguistics, 34342, Istanbul, Turkey

output:

  pdf_document:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    keep_tex: yes
    number_sections: yes
    template: manuscript.latex
    includes:
      in_header: header.tex

  html_document: null

  word_document: null

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: singlespacing
footerdate: yes
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

---

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")

```

```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

```{r init, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(languageR)
library(tidyverse)
library(gdata)
library(MASS)
library(magrittr)
library(ggplot2)

data <- read.csv('results', 
                 header = F, 
                 comment.char = "#", 
                 encoding = "UTF-8" , 
                 col.names = paste0("V",seq_len(11)), 
                 fill = TRUE)

colnames(data) = c("Time", "MD5", "ControllerType", "SentenceNoInStimFile", "Element", "Type", "Item", "Sentence", "Question","Answer", "RT")

FormEntries <- subset(data, ControllerType != "DashedAcceptabilityJudgment" ) %>% drop.levels(); nrow(data)
data %<>% subset(ControllerType == "DashedAcceptabilityJudgment")
age <- dplyr::select(FormEntries, MD5,Sentence,Question) %>% 
  dplyr::filter(Sentence == "age") %>%
  dplyr::select(MD5, Age = Question)
natturk <- dplyr::select(FormEntries, MD5,Sentence,Question) %>% 
  dplyr::filter(Sentence == "natturk") %>%
  dplyr::select(MD5, Natturk = Question) %>%
  mutate(Natturk = Natturk == "male")
dex <- dplyr::select(FormEntries, MD5,Sentence,Question) %>% 
  dplyr::filter(Sentence == "dex") %>%
  dplyr::select(MD5, Dex = Question) %>%
  mutate(Dex = ifelse(Dex == "male", "right-handed", "left-handed"))
FormPart <- dplyr::left_join(age, natturk, by = "MD5") %>% 
  dplyr::left_join(., dex , by = "MD5" ) %>% 
  unique()
lenForm <- length(levels(factor(FormPart$MD5))); lenForm
md5form <- levels(factor(FormPart$MD5))
md5form_subject <- data.frame("MD5" = md5form, "Subject" = sprintf("S[%s]",seq((lenForm))))
FormPart <- merge(FormPart,md5form_subject, all = T)

stopifnot( nrow(data) %% 2 == 0 )
rows_stim <- data[c(T,F),]
rows_resp <- data[c(F,T),]

data <- rows_resp %>% left_join(md5form_subject) %>% dplyr::select(-MD5, -Time, -ControllerType, -Sentence, -Element) %>%
  dplyr::rename(ResponseCorrect=Answer, Response=Question)
data %<>% group_by(Subject) %>% mutate(trial_no = seq(Subject))
data %<>% within({ late_response = (Response == "NULL")
Response[late_response] = NA
ResponseCorrect[late_response] = NA
})

responses <- c(yes="İYİ (P'ye basınız)", no="KÖTÜ (Q'ya basınız)")
stopifnot( all(data$Response %in% responses | is.na(data$Response) ) )

data$ResponseYes <- ifelse(grepl("P",data$Response) , T, 
                           ifelse(grepl("Q",data$Response) , F, NA))

data$condition <- 
  with(data, case_when(Type == "filler" & SentenceNoInStimFile >= 200 ~ "a",
                       Type == "filler" & SentenceNoInStimFile < 200 ~ "b",
                       Type == "condition_a" ~ "a",
                       Type == "condition_b" ~ "b",
                       Type == "condition_c" ~ "c",
                       Type == "condition_d" ~ "d",
                       TRUE ~ as.character(NA)))
data$experiment <- ifelse(data$Type %in% c("condition_a", "condition_b", "condition_c", "condition_d"), "AgrAttr", as.character(data$Type))

data %>% group_by(experiment, condition) %>% 
  summarise(perc_yes = mean(ResponseYes, na.rm = T),
            perc_na = mean(is.na(ResponseYes)))

conditions_info <- data.frame(experiment = c("AgrAttr", "AgrAttr", "AgrAttr", "AgrAttr", "filler", "filler"),
                              condition = c("a","b","c","d", "a","b"),
                              grammatical = c("ungram", "gram", "ungram", "gram", "ungram", "gram"),
                              verb_num = c("sg","pl","sg","pl", "sg","pl"),
                              attracted = c(T,T,F,F,F,F),
                              stringsAsFactors = FALSE) 
data %<>% left_join(conditions_info)
data %<>% subset(Type != "practice") %>% drop.levels()

dataAv <- data %>% 
  group_by(experiment, condition, grammatical, verb_num, attracted) %>%
  summarize(avRT = mean(RT), p_yes = mean(ResponseYes, na.rm = T)) %>% 
  as.data.frame()

dataAv %>% ggplot(aes(grammatical, p_yes, group = paste(experiment,attracted), color = experiment )) + geom_point() + geom_line(aes(linetype =attracted))

data$trial_group <- Hmisc::cut2(data$trial_no, g=5)

dataAvSplit <- data %>% 
  group_by(experiment, condition, grammatical, verb_num, attracted, trial_group) %>%
  summarize(avRT = mean(RT), p_yes = mean(ResponseYes, na.rm = T)) %>% 
  as.data.frame()

plot_split <- dataAvSplit %>% ggplot(aes(grammatical, p_yes, group = paste(experiment,attracted), color = experiment )) + geom_point() + geom_line(aes(linetype =attracted)) + facet_wrap(~trial_group)

dataAvBySubj <- data %>% 
  group_by(Subject, experiment, condition, grammatical, verb_num, attracted) %>%
  summarize(avRT = mean(RT), p_yes = mean(ResponseYes, na.rm = T), N = sum(!is.na(ResponseYes))  )  %>% 
  as.data.frame()
head(dataAvBySubj,10)

dataAvBySubjWide = dataAvBySubj %>% mutate(expcond = paste(experiment,condition, sep="_")) %>% dplyr::select(-experiment, -condition, -avRT, -N, -grammatical, -verb_num, -attracted) %>% tidyr::spread(expcond, p_yes)
dataAvBySubjWide %<>% mutate(delta_dc = AgrAttr_d - AgrAttr_c)
bad_subjects = subset(dataAvBySubjWide, delta_dc <= 0.25 ) %>% .$Subject

dataAvBySubj %>% subset(Subject %in% bad_subjects) %>% ggplot(aes(grammatical, p_yes, group = paste(experiment,attracted), color = experiment )) + geom_point() + geom_line(aes(linetype =attracted)) + facet_wrap(~Subject)

dataAvClean <- data %>% subset(!Subject %in% bad_subjects)%>% 
  group_by(experiment, condition, grammatical, verb_num, attracted) %>%
  summarize(avRT = mean(RT), p_yes = mean(ResponseYes, na.rm = T)) %>% 
  as.data.frame()

dataAvClean <- slice(dataAvClean, 1:4)

plot_average <- dataAvClean %>% ggplot(aes(grammatical, p_yes, group = paste(experiment,attracted), color = experiment )) + geom_point() + geom_line(aes(linetype =attracted)) #+ facet_wrap(~experiment, scales = "free_y")

```

\vspace{2mm}\hrule

In this paper, our objective is to explore the source of agreement attraction effects in Turkish found by @Lago. The interpretation of their finding is complicated by the fact that in their possessive form all head nouns in their stimuli are morphologically ambiguous between possessive and accusative. Due to the mismatch in the frequency of usage of accusative and genitive, which is used as a attractor in this experiment, as a controller, @Lago's findings could potentially be explained with occasional shallow processing. We have replicated @Lago's experiment with unambigious head nouns, and we have found the same agreement attraction effects. However, this may mean that participants may engage in even more shallow processing than we imagined. We postulate that participants solely check for the presence of a *-lAr* morpheme, while largely disregarding the remainder of the sentence. We utilize an experiment design in which we use a relative clause with or without plural agreement on the verb in place of the genitive possessor since both verbal and nominal plural morphemes are the same in Turkish. If participants did, in fact, adopt a superficial processing strategy, we would expect to find a number agreement attraction effect similar in magnitude to @Lago's. However, if number attraction is not an artifact of shallow processing strategies, we should expect to find no number attraction in this experiment.

\vspace{3mm}\hrule

*Keywords*: Turkish, agreement attraction, task effect, shallow processing

\bleft

---
nocite: | 
  @woolford
...

# INTRODUCTION

Attraction errors in production and comprehension of subject-verb agrrement, in which a verb does not agree with the grammatical agreement controller, but with a potential attractors, have been the focal point of many research for quite a long time. In fact, it is still a widely researched are in psycholinguistic studies. Despite the thorough research that has been carried out, studies that have been conducted on agreement attraction in Turkish have been extremely few. In fact, @Lago has been the only study to look into this phenomenon in Turkish. @Lago's study make use of genitive-possessive structures in the subject position, in which the possessive-marked noun is the head of the noun phrase which acts as the grammatical agreement controller, and the genitive noun serves as a potential attractor. In a speeded acceptability judgment study, they found a significant effect of number agreement attraction. However, the interpretation of their finding may be a result of a non-subjecthood cues originating from their use of morphologically ambiguous forms of possessive. In their possessive form all head nouns in their stimuli are ambigiuous between possessive and accusative.

In Turkish, accusative number agreement controllers are extremely rare, while genitive agreement controllers are very frequent. Thus, @Lago's finding could possibly be explained by occasional shallow processing. When all syntactic relations in the sentence were processed fully, the possessive noun should have been identified as the controller. Meanwhile the genitive noun may sometimes have been erroneously identified as the controller during shallow processing, because genitives are more likely to act as agreement controllers than accusatives. A second alternative explanation of @Lago's finding may be the fact that participants may engage in even more shallow processing than outlined above. Participants may have erroneously responded *Yes* on some trials due to (i) the presence of a plural morpheme on a noun (attractor or controller) and (ii) the presence of a plural agreement morpheme on the verb. We speculate that on such trials, participants would have simply tried to check for the presence of such morphemes, while largely disregarding the remainder of the sentence.

We first replicated @Lago's experiment with unambiuous head nouns. To this end, we have revised the items that were used previously, to avoid morphological ambiguity between possessive and accusative forms. The effect found by @Lago replicates with unambiguous nouns as discused in \S 2.

\S 3 discussses the alternative account that posits even more shallow processing as mentioned above and offers a pre-registered experiment using RC attractors with possible outcomes and their indications. Since both nominal and verbal plural morphemes in Turkish take on the same form (*-ler* or *-lar* depending on the phonological environment), we can test this possibility with an experiment, in which we use relative clause with or without plural agreement on the verb in the place of the genitive possessors. An agreement attraction effect similar in magnitude to @Lago's would mean that participants use aforementioned strategies.

\S 4 presents a discussion revolving around possible 
Lastly, \S 5 offers a brief conclusion and presents topics for future researches.

# REPLICATON OF @Lago

In their study, @Lago investigated the comprehension of subject-verb agreement in Turkish-German bilinguals and Turkish monolinguals. They used speeded acceptability judments for the effects of number attraction in Turkish. Their sentences makes us of genitive-possessive constructions in the subject position, where the genitive is the attractor and the possessive is the head noun. They have manipulated the grammaticality of the sentence by changing the plural morphology of the verb, and they also manipulated the plurality of the attractor noun. In grammatical conditions, subject and the verb both bears the singular morphology with no overt morpheme. Moreover, in the ungrammatical conditions the verb bears the overt *-lAr* morpheme whereas the subject is still singular as exemplified below.

\begin{exe}
\ex
\begin{xlist}
\ex \underline{Grammatical, SG attractor} \label{lago1}
\gll \c{S}ark{\i}c{\i}-n{\i}n vokalist-i sahne-de s\"{u}rekli z{\i}pla-d{\i}\\
singer-\textsc{gen} vocalist-\textsc{poss} stage-\textsc{loc} non-stop jump-\textsc{pst}-$\varnothing$\\
\glt The singer's backup vocalist jumped on the stage non-stop.

\ex \underline{Grammatical, PL attractor} \label{lago2}
\gll \c{S}ark{\i}c{\i}-lar-{\i}n vokalist-i sahne-de s\"{u}rekli z{\i}pla-d{\i}\\
singer-\textsc{pl}-\textsc{gen} vocalist-\textsc{poss} stage-\textsc{loc} non-stop jump-\textsc{pst}-$\varnothing$\\
\glt The singers' backup vocalist jumped on the stage non-stop.

\ex \underline{Ungrammatical, PL attractor} \label{lago3}
\gll \c{S}ark{\i}c{\i}-lar-{\i}n vokalist-i sahne-de s\"{u}rekli z{\i}pla-d{\i}-lar.\\
singer-\textsc{pl}-\textsc{gen} vocalist-\textsc{poss} stage-\textsc{loc} non-stop jump-\textsc{pst}-\textsc{3Pl}\\
\glt The singers' backup vocalist jumped on the stage non-stop.

\ex \underline{Ungrammatical, SG attractor} \label{lago4}
\gll \c{S}ark{\i}c{\i}-n{\i}n vokalist-i sahne-de s\"{u}rekli z{\i}pla-d{\i}-lar.\\
singer-\textsc{gen} vocalist-\textsc{poss} stage-\textsc{loc} non-stop jump-\textsc{pst}-\textsc{3Pl}\\
\glt The singer's backup vocalist jumped on the stage non-stop.
\end{xlist}
\end{exe}

They have found a significant effect of number attraction in Turkish ranging between 11%--15% across monolinguals. As seen from the results of statistical analysis in \textsc{Table} (\ref{lagomodel}), acceptability judgments showed an immense effect of grammaticality, and there is also interaction between grammaticality and attractor number which indicates an number attraction effect.

\begin{table}[!hbt]
\centering
\begin{tabular}{llcccc}
\hline
    &                                        & \multicolumn{4}{l}{Monolingual Speakers}                                               \\ \cline{3-6} 
    &                                        & $\beta$        & SE            & \emph{z} & \emph{p} \\ \hline
\multicolumn{2}{l}{\textbf{Attraction Task}} &                &               &                           &                           \\
    & Grammaticality                         & \textbf{-5.51} & \textbf{0.33} & \textbf{-16.69}           & \textbf{.000}             \\
    & Attractor Number                       & 0.14           & 0.25          & 0.57                      & .571                      \\
    & Grammaticality x Attractor Number      & \textbf{1.69}  & \textbf{0.53} & \textbf{3.19}             & \textbf{.001}             \\
    & Attractor Number: Ungram conditions    & \textbf{0.94}  & \textbf{0.26} & \textbf{3.68}             & \textbf{.000}             \\
    & Attractor Number: Gram conditions      & -0.79          & 0.52          & -1.51                     & .131                      \\ \hline
\end{tabular}
\caption{Model results for the judgments of monolingual cited from @Lago.}
\label{lagomodel}
\end{table}

According to @Lago's stipulation, Turkish genitive case does not provide a strong cue against subjecthood since it is extremely common to see genitive marked subjects in Turkish embedded clauses, as in example (\ref{gensubj}). Unlike English, genitive case is compatible with the subjecthood in English. Thus, @Lago attributes the robust agreement attraction effects to the case information.

\begin{exe}
\ex \label{gensubj}
\gll k\"{o}y-\"{u} bir haydut-un bas-t{\i}\u{g}-{\i}n-{\i} duy-du-m.\\
village-\textsc{acc} a bandit-\textsc{gen} raid-\textsc{nmlz}-\textsc{3Sg}-\textsc{acc} hear-\textsc{pst}-\textsc{1Sg}\\
\glt I heard that a (certain) robber raided the village. (Adapted from Woolford (2009))
\end{exe}

However, our first hypothesis was the possibility of participants engaging in a shallow processing, which ends up in a situation where not only genitive case but also possessive case plays a significant role in delivering the case information. None of the experimental items in @Lago has a head noun which ends with a vowel; thus, all of the possessive markers are morphologically and phonologically ambiguous between the accusative case and the possessive. Unlike genitive case, the accusative case in Turkish is extremely rare to see on the head noun of the subject, which only occurs in the raising predicates as in example (\ref{dsm}).

\begin{exe}
\ex \label{dsm}
\gll Ben sen-i git-ti-n san-d{\i}-m\\
I you-\textsc{acc} go-\textsc{pst}-\textsc{2Sg} suppose-\textsc{pst}-\textsc{1Sg}\\
\glt I thought you were gone.
\end{exe}

In order to test this hypothesis, we have replicated the @Lago study with modified items, where the possessive is not ambiguous with the accusative case. We have tried to be as faithful as possible to the original sentences while trying to make sentences as plauisable as possible at the same time. We keep the semantic relation between the head nound and the controller the same with   A set of sentences are exemplified below.

\begin{exe}
\ex
\begin{xlist}
\ex \underline{Grammatical, SG attractor} \label{exp1}
\gll Komedyen-in yard{\i}mc{\i}-s{\i} poyraz-dan dolayi \"{u}\c{s}\"{u}-d\"{u}.\\
comedian-\textsc{gen} helper-\textsc{poss} northeaster-\textsc{abl} because.of feel.chilly-\textsc{pst}-$\varnothing$\\
\glt Because of the northeaster, comedian's assistant felt chilly.

\ex \underline{Grammatical, PL attractor} \label{exp2}
\gll Komedyen-ler-in yard{\i}mc{\i}-s{\i} poyraz-dan dolayi \"{u}\c{s}\"{u}-d\"{u}.\\
comedian-\textsc{pl}-\textsc{gen} helper-\textsc{poss} northeaster-\textsc{abl} because.of feel.chilly-\textsc{pst}-$\varnothing$\\
\glt Because of the northeaster, comedians' assistant felt chilly.

\ex \underline{Ungrammatical, PL attractor} \label{exp3}
\gll Komedyen-ler-in yard{\i}mc{\i}-s{\i} poyraz-dan dolayi \"{u}\c{s}\"{u}-d\"{u}-ler.\\
comedian-\textsc{pl}-\textsc{gen} helper-\textsc{poss} northeaster-\textsc{abl} because.of feel.chilly-\textsc{pst}--\textsc{pl}\\
\glt Because of the northeaster, comedians' assistant felt chilly.

\ex \underline{Ungrammatical, SG attractor} \label{exp4}
\gll Komedyen-in yard{\i}mc{\i}-s{\i} poyraz-dan dolayi \"{u}\c{s}\"{u}-d\"{u}-ler.\\
comedian-\textsc{gen} helper-\textsc{poss} northeaster-\textsc{abl} because.of feel.chilly-\textsc{pst}-\textsc{pl}\\
\glt Because of the northeaster, comedian's assistant felt chilly.

\end{xlist}
\end{exe}

As seen in the examples unlike @Lago's experimental items, all of our items bear the *-sI* possessive marker instead of the ambiguous *-I* one. As for the filler items, we could not used or adapt from the original experimentsince the fillers of the original study were not online. We have used two types filler sentences: a grammatical ones where the verb bears plural agreement (\ref{fillera}), and an ungrammatical one where verb does not bear plural agreement (\ref{fillerb}). We wanted to nullify a possible strategy by the participants where they associate the sentence-final morpheme directly with the acceptability, and disregard other elements in the sentences when answering questions.

\begin{exe}
\ex
\begin{xlist}
\ex \label{fillera}
\gll Adam-{\i}n anne-si fena-la\c{s}-{\i}nca inek kurban et-ti-ler.\\
man-\textsc{gen} mother-\textsc{poss} bad-\textsc{vrb}-\textsc{cvb} cow sacrifice do-\textsc{pst}-\textsc{pl}\\
\glt When his mother got ill, (they) sacrificed a cow.
\ex \label{fillerb}
\gll *Pizzac{\i}-n{\i}n kurye-si t\"{o}kezle-yince sos-lar yer-e sa\c{c}-t{\i}.\\
pizzaria-\textsc{gen} courier-\textsc{poss} trip-\textsc{cvb} sauce-\textsc{pl} floor-\textsc{dat} scatter-\textsc{pst}\\
\glt Intended: When the pizza boy tripped, sauces scattered around. 
\end{xlist}
\end{exe}

All of our data, experimental materials, our experiment design, and as well as our fillers can be found at the Center for Open Science Framework website (https://osf.io/).

## Participants and Procedure

One hundred seven Turkish speakers with mean age of X were recruited from the Bogazici University in {\.I}stanbul. We did not collect participants' knowledge of other languages; however, we ensured that Turkish is their native language and they predominantly used in their daily lives. In the experiments, participants are asked to judge the experimental and filler sentences whether or not they sound acceptable in Turkish. All of the sentences were presented one word at a time in the center area of the screen for 500 ms per word unlike @Lago and @Wagers2009, in which researches used 300 ms visual presentation per word. Experiment was run on a web-based platform Ibex Farm, and all of the documentation can be found in our osf and github page.

Before the experiment, they have been instructed to give accurate and quick answers with their own intuition, and they were also notified about the time limit for answering. At the start of the experiment they were given 4 practice items with feedbacks.

What to write about analyze?

## Results and Discussion

In \textsc{Figure} (\ref{fig:replication}), y-axis shows proportion of "acceptable" answers, and x-axis indicates whether or not the sentences in that group is grammatical. Moreover, linetypes indicates an attractor noun with overt plural morphology. As seen from the figure, `r round(100*dataAvClean[1,7])`% of the sentences with plural attractors and singular verb were accepted by the participants following the findings of @Lago. We also see a number attraction rate of `r round(100*(dataAvClean[1,7]-dataAvClean[3,7]))`% which was already shown in @Lago with close results.

In \textsc{Table} (\ref{tab:replication}), ............

\begin{figure}[H]
  \begin{floatrow}
    \ffigbox{%
```{r, fig.align = "right", echo = F}
print{}
```
    }{\caption{A figure}}

    \btabbox{%
```{r, fig.align = "right", echo = F}
knitr::kable(head(mtcars[,1:3]), format = "latex")
```
    }{\caption{A table}}
  \end{floatrow}
\end{figure}

# EXPERIMENT 2: RC ATTRACTOR
## Motivation
## Items and Fillers
## Expectations
## Possible Outcomes and Discussion

# DISCUSSION

Discuss.

# CONCLUSIONS

Intro and repeat.

# ACKNOWLEDGEMENTS
We used the statistical language `R` `r citep(citation())` for all our analyses. These were implemented in dynamic rmarkdown documents using `knitr` `r citep(citation("knitr"))` and `rmarkdown` `r citep(citation("rmarkdown"))` packages. All graphs have been done with `ggplot` `r citep(citation("ggplot2"))`. Thank the Mine and Kadir for their student from courses.

# REFERENCES

```{r write_citations, cache=FALSE, include=FALSE}
write.bibtex(file="knitcitations.bib")
```

<div id = "refs"></div>

\eleft

\clearpage

\listoftables

\newpage

```{r Table1, results='asis', echo=FALSE, cache=FALSE}
kable(head(iris), caption = "A glimpse of the famous *Iris* dataset.")
```

\newpage

```{r Table2, results='asis', echo=FALSE, cache=FALSE}
kable(mtcars[10:16, ], caption = "Now a subset of mtcars dataset.")
```

\clearpage

\listoffigures

\newpage

```{r Fig1, echo=FALSE, fig.cap="Just my first figure with a very fantastic caption.", cache=FALSE}
x <- rnorm(100)
y <- jitter(x, 1000)
plot(x, y)
```

\newpage

\blandscape

```{r Fig2, echo=FALSE, fig.cap="Second figure in landscape format.", cache=FALSE}
a <- sort(rnorm(100))
b <- c(rep("Group Small", 35), rep("Group Big", 65))
boxplot(a ~ b)
```

\elandscape

\clearpage

```{r sessioninfo, echo = FALSE, eval = FALSE}
# set eval = FALSE if you don't want this info (useful for reproducibility) to appear 
sessionInfo()
```