### 2019-12-17 agremeent_draft_dec17/settings.tex

\let\B\relax %to resolve a conflict in the definition of these commands between xyling and xunicode (the latter called by fontspec, called by charis)

\let\T\relax

\usepackage{xyling} %for trees; the use of xyling with the CharisSIL font produces poor results in the branches. This problem does not arise with the packages qtree or forest.
\usepackage[linguistics]{forest} %for nice trees!

\usepackage{leipzig}
\makeglossaries
\renewcommand{\glossarypreamble}{%
  Abbreviations are as follows:{\ }}

\usepackage{linguex}
\usepackage{hyperref}