### 2019-12-23 agremeent_draft_dec17/settings.tex
models are run,use lfs for big files

\let\B\relax %to resolve a conflict in the definition of these commands between xyling and xunicode (the latter called by fontspec, called by charis)

\let\T\relax

\usepackage{xyling} %for trees; the use of xyling with the CharisSIL font produces poor results in the branches. This problem does not arise with the packages qtree or forest.
\usepackage[linguistics]{forest} %for nice trees!

\usepackage[block]{leipzig}
% input glosses
\makeglossaries

\newleipzig{aor}{aor}{aorist}

\usepackage{linguex}
\usepackage{hyperref}